---
title: "EDA Project Diamond Prices"
author: "Francia Moscoso"
date: "March 2, 2016"
output: html_document
---
**Loading Libraries and 'diamonds' data set**
```{r LoadLib, message=F, warning=F}
library(ggplot2)
library(gridExtra)
library(dplyr)
library(lubridate)
```

To set up working directory: setwd("~/SprintboardProject")

Use function install.packages("LibraryName")  to install data and time package


To silence the execution of the library function eg.library(dplyr), use statement: {r message=F, warning=F}
```{r loadDiamonds, comment="", echo=TRUE}
data("diamonds")
str(diamonds)
```


**How many observations are in the data set?**
53940 obs.

**How many variables are in the data set?**
10 variables

**How many ordered factors are in the data set?**
2

**What letter represents the best color for a diamond?**
Levels: D < E < F < G < H < I < J

**What letter represents the best color for a diamond?**
According with the help file: color: diamond colour, from J (worst) to D (best)


**Create a histogram of the price of all the diamonds in the diamond data set**
Long-tailed distribution

```{r structure}
hist(diamonds$price, main="Histogram of Diamond Prices", plot = T,breaks =40,col="yellow")
```

** How many diamonds cost less than $500?
```{r lessTHan500, comment="", echo=FALSE}
filter(diamonds, price < 500)
```
** How many diamonds cost less than $250?**
```{r comment="", echo=FALSE}
filter(diamonds, price < 250)
```

**How many diamonds cost $15,000 or more?**
```{r comment="", echo=FALSE}
filter(diamonds, price >=  15000)
```





**Break Out the histogram of diamond prices by cut. You should have five histograms in separate panels on your resulting plot.**

```{r DiamondFair}
hist(filter(diamonds,diamonds$cut == levels(diamonds$cut)[1])$price,
     main="Histogram of Fair Diamond Prices", 
     plot = T,col="yellow", 
     breaks=c(200, seq(300,18800, 100)),
     xlim=c(0, 20000),  ylim=c(0,70))
```


```{r DiamondGood}
hist(filter(diamonds,diamonds$cut == levels(diamonds$cut)[2])$price,
     main="Histogram of Good Diamond Prices", 
     plot = T,col="yellow", 
     breaks=c(200, seq(300,18800, 100)),
     xlim=c(0, 20000),  ylim=c(0,300))
```


```{r DiamondVeryGood}
hist(filter(diamonds,diamonds$cut == levels(diamonds$cut)[3])$price,
     main="Histogram of Very Good Diamond Prices", 
     plot = T,col="yellow", 
     breaks=c(200, seq(300,20000, 100)),
     xlim=c(0, 20000),  ylim=c(0,800))
```


```{r DiamondPremiun}
hist(filter(diamonds,diamonds$cut == levels(diamonds$cut)[4])$price,
     main="Histogram of Premium Diamond Prices", 
     plot = T,col="yellow", 
     breaks=c(200, seq(300,20000, 100)),
     xlim=c(0, 20000),  ylim=c(0,800))
```


```{r DiamondIdeal}
hist(filter(diamonds,diamonds$cut == levels(diamonds$cut)[5])$price,
     main="Histogram of IDeal Diamond Prices", 
     plot = T,col="yellow", 
     breaks=c(200, seq(300,20000, 100)),
     xlim=c(0, 20000),  ylim=c(0,800))
```



Indeed, we can see that the prices for diamonds are heavily skewed, but when put on a log10 scale seem much better behaved (i.e., closer to the bell curve of a normal distribution).
```{r}
 qplot(price, data = diamonds, binwidth = 0.01) + scale_x_log10() +
   theme_bw() +
   ggtitle("Price")
```

The command scales="free_y" allows the y-axis to be different for each histogram.  Also, see the difference between binwidth=10 vs binwidth=80
```{r }
qplot (x = price, data = diamonds, binwidth=10) + facet_wrap(~cut, scales="free_y")

qplot (x = price, data = diamonds, binwidth=80) + facet_wrap(~cut, scales="free_y") 
```
The command scales="free_y" allows the y-axis to be different for each histogram.  Also, see the difference between binwidth=10 vs binwidth=80


**Summarize data to get max, min, medium, lowest by Diamond Cut.**<br>
1. Which cut has the highest priced diamond? Premium
2. Which cut has the lowest priced diamond? Very Good, Premium and Ideal
3. WHich cut has the lowest median price? Ideal

```{r comment="", echo=FALSE}
diamondCut <- group_by(diamonds, cut)
diamondByCut <- summarise(diamondCut, mean_price=mean(price),
                median_price = median(price), min_price=min(price),
                max_price = max(price))
diamondByCut
```

**Create a Histogram of price per carat and facet it by cut.  You can make adjustmentsto the code from the previous exercise to get started. Adjust the bin width and transform the scale of the x-axis using log10**

```{r}
ggplot(aes(x = (price / carat)), data = diamonds) + 
geom_histogram(binwidth = 0.10,color = "black", fill = "steelblue") +
scale_x_log10() +
facet_wrap( ~ cut, scales = "free") +
labs(x = "Price per Carat", y = "Number of Diamonds")
```

Price vs. Carat - Vertical 
```{r}
qplot(log(price), carat, data = diamonds, facets = cut ~ .,
geom =c("point","smooth"))
```

Price vs. Carat - Horizontal
```{r}
qplot(log(price), carat, data = diamonds,
geom=c("point","smooth"),facets = . ~ cut) 
```


Dividing Carat in quantiles
```{r comment="", echo=FALSE}
cutpoints <- quantile(diamonds$carat, seq(0,1, length = 4), na.rm = TRUE)
cutpoints
diamonds$caratVal <- cut(diamonds$carat, cutpoints)
levels(diamonds$caratVal)

qplot(log(price), data=diamonds,  facets = caratVal ~ cut, 
     geom =c("histogram"), binwidth=0.10) 
#    geom =c("histogram","smooth"))  # The smoth gives error

```



**Investigate the price of diamonds using box plots, numerical summaries, and one of the following categorical variables: cut, clarity, or color**
```{r comment="", echo=FALSE}
qplot(x=cut, y=price,data=diamonds,geom='boxplot',fill=cut) +
      coord_cartesian(ylim=c(300,6500))

diamonds%>%group_by(clarity)%>%
   summarise('maximum price'=max(price),
   'minimun price'=min(price))
   
```


**Investigate the price per carat of diamonds using boxplots.**
```{r comment="", echo=FALSE}
p1 = qplot(data=diamonds, x=cut, y=price,
geom='boxplot', color=cut) +
coord_cartesian(ylim=c(0,7000))
p2 = qplot(data=diamonds, x=clarity, y=price,
geom='boxplot', color=clarity) +
coord_cartesian(ylim=c(0,7000))
p3 = qplot(data=diamonds, x=color, y=price,
geom='boxplot', color=color) +
coord_cartesian(ylim=c(0,9000))
grid.arrange(p1, p2, p3)
```


**What is the price range for the middle 50% of diamonds with color D?**<br>
First Quartile (25%): 911
Third Quartile (75%): 4214
```{r comment="", echo=FALSE}
DColor = filter(diamonds, diamonds$color == "D")
quantile(DColor$price, c(0.25, 0.5, 0.75), type = 1)

fivenum(DColor$price)

quantile(filter(diamonds, diamonds$color == "D")$price,
          c(0.25, 0.5, 0.75), type = 1)
```


**What is the price range for the middle 50% of diamonds with color J?**<br>
First Quartile (25%): 1860
Third Quartile (75%): 7695
```{r comment="", echo=FALSE}
JColor = filter(diamonds, diamonds$color == "J")
quantile(JColor$price, c(0.25, 0.5, 0.75), type=1)

fivenum(JColor$price)

quantile(filter(diamonds, diamonds$color == "J")$price,
          c(0.25, 0.5, 0.75), type = 1)
```



**What is the IQR for diamonds with the best color?**(D is the best color) <br>
3302
```{r comment="", echo=FALSE}
IQR(DColor$price)
```


**What is the IQR for diamonds with the worst color?**(D is the best color)<br>
5834
```{r comment="", echo=FALSE}
IQR(JColor$price)
```



**Investigate the weight of the diamonds (carat) using a frequency poligon.  Use different bin widths to see how the frequency polygon changes.  What carat size has a count greater than 2000?**
**0.1; 0.3; 0.8; 1.01; 1.6; 2.0; 3.0; 5.0** Looks like none
```{r}
qplot(x = diamonds$carat,
      data = diamonds,
      binwidth = 0.1,
      xlab = 'Carats',
      ylab = 'Count',
      geom = 'freqpoly') +
      scale_x_continuous(breaks=1:5) + scale_y_continuous(limits=c(0, 12000))
```

**Zooming the graphic to display x-axis: 0.1-2 and y-axis: 0-3000 to investigate if they have a count greater than 2000**

```{r}
qplot(x = diamonds$carat,
      data = diamonds,
      binwidth = 0.2,
      xlab = 'Carats',
      ylab = 'Count',
      geom = 'freqpoly') +
      scale_x_continuous(limits = c(0,2), breaks = seq(0,2,0.2) ) +
      scale_y_continuous(limits = c(0,5000), breaks = seq(0,5000,250))
```

**Zooming - Another look x-axis between 0 and 2** 
```{r,
qplot(x = diamonds$carat,
      data = diamonds,
      binwidth = 0.1,
      xlab = 'Carats',
      ylab = 'Count',
      geom = 'freqpoly') +
      scale_x_continuous(limits = c(0,2), breaks = seq(0,2,0.1) ) +
      scale_y_continuous(limits = c(0,5000), breaks = seq(0,5000,250))  
```

**Look how it changes if we increase to 0.5 the binwidth (...seq(0.1,2.1,0.5))**
```{r}
qplot(x = diamonds$carat,
      data = diamonds, 
      binwidth = 0.1,
      xlab = 'Carats',
      ylab = 'Count',
      geom = 'freqpoly') +
      scale_x_continuous(limits = c(0.1,2.1), breaks = seq(0.1,2.1,0.5) ) +
      scale_y_continuous(limits = c(0,2000), breaks = seq(0,2000,100))
```


**Another one**
**Look how there is a binwidth and no breaks in the scale_x_continuous.**
**In the next graphic (not the one below) you will see the a break in the scale_x_continuous**
```{r}
qplot(x = diamonds$carat,
      data = diamonds,
      binwidth = 0.01,
      xlab = 'Carats',
      ylab = 'Count',
      geom = 'freqpoly') +
      scale_x_continuous(limits = c(0,3)) +
      scale_y_continuous(limits = c(0,2000), breaks = seq(0,2000,1000)) 
```


**Compare the following graphic with the previous one. Look how this time there is a tick for every 0.3. Also, we are omitting the break for the y axis**
```{r}
qplot(x = diamonds$carat,
      data = diamonds,
      binwidth = 0.1,
      xlab = 'Carats',
      ylab = 'Count',
      geom = 'freqpoly') +
      scale_x_continuous(limits = c(0,3), breaks = seq(0,3, 0.3) )  +
      scale_y_continuous(limits = c(0,2000)) 
```




